{\plant}: A package for modelling forest trait ecology \& evolution

### 2015-11-09 ms.tex
Rename EBT to SCM
This changed is motivated by the fact we are no longer using the EBT method, but rather "characteristic method".

Acronym stands for "Solver Chacteristic Method"

Corresponds to changes implemented in traitecoevo/plant/commit/b2ab54283b95e186c511720d71672ac2f0e57607
Resolve issue around naming and symbols for size distribution
Suggestion from Ulf. Previous approach was misleading, because N(x,H,a)
would imply density of plants per unit height per unit trait per unit age,
whereas as we were only recording density per unit height.

Instead use N(H|x,a)

When modelling a patch of competing plants, the focus is on how the size-density-distribution \(N(x, H, a)\) changes with patch age \(a\), with the latter defined as the time that has passed since the last disturbance. This distribution describes the density \(N\) of
individuals with traits \(x\) and height \(H\) within a patch of age \(a\).
The density \(N\) is measured as individuals per unit height and per unit ground area.
Modelling the dynamics of \(N(x, H, a)\) requires that both the initial size-density-distribution and the inflow of new recruits be specified. We have primarily
been interested in the dynamics of a patch recovering from disturbance and have
therefore started with an empty patch and a constant flow of seeds from a global
seed rain (Fig. \ref{fig:schematic}b). A (non-trivial) extension would result from allowing some plants to survive disturbance \citep[following][]{Kohyama-1993}.

{\plant} offers two methods for modelling the dynamics of \(N(x,
H, a)\) in a competing population. In the first stochastic mode, users
specify an average seed rain and patch size, and {\plant} then generates a
vector of seed arrival events and simulates the stochastic
development of the resulting finite-sized population.

@@

Our approach for numerically solving deterministic size-structured population dynamics is
based on the characteristic method \citep{Angulo-2004}. This method describes the development of \(N\) by approximating the
density function along a collection of individual growth trajectories spanning the size
spectrum. Following a disturbance, a series of such trajectories are introduced
into each patch. These trajectories then change according to the growth
of the corresponding individuals, conditioned on their survival
(Fig. \ref{fig:patch}a). Meanwhile, growth and mortality combine
to alter the density of individuals along these trajectories
(Fig. \ref{fig:patch}b). The characteristic method is similar to the Escalator Boxcar Train (\textsc{ebt}) method
\citep{Deroos-1988, Deroos-1997} used by \citet{Falster-2011} and
\citet{Falster-2015}, but not identical. An advantage of the characteristic method is that it allows direct approximation of the size-density-distribution. For the same reason, however, the characteristic method may be less suitable for resolving size-density-distributions in models where those distributions naturally develop sharp peaks (which happens, e.g., when plants completely cease to grow at a given height).

@@

Such dynamic equilibria arising in \textsc{TSPMs} potentially resolve several
challenges faced by unstructured models. First, by scaling up from
patches to a metapopulation, we close the demographic feedback loop and
create a self-regulating system. {\plant} yields as outputs the equilibrium seed rains and size-density-distributions for each considered species, based on the combined effects of the model's physiological rules, species traits, competitive interactions
and disturbance regime.

>>>>

When modelling a patch of competing plants, the focus is on how the size distribution \(N(H | x, a)\) changes with patch age \(a\), with the latter defined as the time that has passed since the last disturbance. This distribution describes the density \(N\) of
individuals with height \(H\) for given traits \(x\) and patch age \(a\).
The density \(N\) is measured as individuals per unit height and per unit ground area.
Modelling the dynamics of \(N(H | x, a)\) requires that both the initial size distribution and the inflow of new recruits be specified. We have primarily
been interested in the dynamics of a patch recovering from disturbance and have
therefore started with an empty patch and a constant flow of seeds from a global
seed rain (Fig. \ref{fig:schematic}b). A (non-trivial) extension would result from allowing some plants to survive disturbance \citep[following][]{Kohyama-1993}.

{\plant} offers two methods for modelling the dynamics of \(N(H | x, a)\)
in a competing population. In the first stochastic mode, users
specify an average seed rain and patch size, and {\plant} then generates a
vector of seed arrival events and simulates the stochastic
development of the resulting finite-sized population.

@@

Our approach for numerically solving deterministic size-structured population dynamics is
based on the characteristic method \citep{Angulo-2004}. This method describes the development of \(N\) by approximating the
density function along a collection of individual growth trajectories spanning the size
spectrum. Following a disturbance, a series of such trajectories are introduced
into each patch. These trajectories then change according to the growth
of the corresponding individuals, conditioned on their survival
(Fig. \ref{fig:patch}a). Meanwhile, growth and mortality combine
to alter the density of individuals along these trajectories
(Fig. \ref{fig:patch}b). The characteristic method is similar to the Escalator Boxcar Train method
\citep{Deroos-1988, Deroos-1997} used by \citet{Falster-2011} and
\citet{Falster-2015}, but not identical. An advantage of the characteristic method is that it allows direct approximation of the size distribution. For the same reason, however, the characteristic method may be less suitable for resolving size distributions in models where those distributions naturally develop sharp peaks (which happens, e.g., when plants completely cease to grow at a given height).

@@

Such dynamic equilibria arising in \textsc{TSPMs} potentially resolve several
challenges faced by unstructured models. First, by scaling up from
patches to a metapopulation, we close the demographic feedback loop and
create a self-regulating system. {\plant} yields as outputs the equilibrium seed rains and size distributions for each considered species, based on the combined effects of the model's physiological rules, species traits, competitive interactions
and disturbance regime.
